---
title: "GAM Assumption Checks"
author: "Michael W Belitz"
date: "2024-03-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(lubridate)
library(DHARMa)
library(mgcv)
# read in adult data
adult_df <- read.csv("data/field_observations/surveyDateSheet.csv") %>% 
  mutate(doy = yday(mdy(eventDate)),
         year = year(mdy(eventDate))) %>% 
  mutate(doy2 = if_else(condition = year == 2019,
                        true = doy,
                        false = doy + 365))

adult_gb <- adult_df %>% 
  group_by(doy2, location) %>% 
  summarise(macroMoths = sum(macroMoths),
            microMoths = sum(microMoths))

# read in lunar illumination
lunar.phase <- read.csv("data/urbanStressors/lunarIllumination.csv") %>% 
  mutate(Date = mdy(Date)) %>% 
  mutate(doy = yday(Date),
         year = year(Date)) %>% 
  mutate(doy2 = if_else(condition = year == 2019,
                        true = doy,
                        false = doy + 365))

adult_gb <- left_join(adult_gb, lunar.phase, by = "doy2")

# get adult df for each location
auca_adult_df <- filter(adult_gb, location == "AUCA")
rist_adult_df <- filter(adult_gb, location == "RIST")
prcr_adult_df <- filter(adult_gb, location == "PRCR")

bowa_adult_df <- filter(adult_gb, location == "BOWA")
biva_adult_df <- filter(adult_gb, location == "BIVA")
demi_adult_df <- filter(adult_gb, location == "DEMI")

cofr_adult_df <- filter(adult_gb, location == "COFR")
baca_adult_df <- filter(adult_gb, location == "BACA")
joma_adult_df <- filter(adult_gb, location == "JOMA")


generateDHARMa <- function(site_df, distributionFamily){
  
  g <- gam(macroMoths ~ s(doy, k = 6, bs = "cr") + s(lunar.phase, k = 3, bs = "cr"),  
         data = site_df, family = distributionFamily)

res = DHARMa::simulateResiduals(g)
resPlot <- plot(res)

return(resPlot)
  
}

generateDHARMaMicro <- function(site_df, distributionFamily){
  
  g <- gam(microMoths ~ s(doy, k = 6, bs = "cr") + s(lunar.phase, k = 3, bs = "cr"),  
         data = site_df, family = distributionFamily)

res = DHARMa::simulateResiduals(g)
resPlot <- plot(res)

return(resPlot)
  
}

```

## Pooled Macro-moth checks

### First negative binomial distributions

Site: AUCA
```{r macroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = auca_adult_df, distributionFamily = negbin(3))

```

Site: PRCR
```{r prcrmacroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = prcr_adult_df, distributionFamily = negbin(3))

```

Site: Rist
```{r ristmacroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = rist_adult_df, distributionFamily = negbin(3))

```

Site: Demi
```{r demimacroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = demi_adult_df, distributionFamily = negbin(3))

```

Site: Bowa
```{r bowamacroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = bowa_adult_df, distributionFamily = negbin(3))

```

Site: Biva
```{r bivamacroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = biva_adult_df, distributionFamily = negbin(3))

```

Site: Baca
```{r bacamacroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = baca_adult_df, distributionFamily = negbin(3))

```

Site: Cofr
```{r cofrmacroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = cofr_adult_df, distributionFamily = negbin(3))

```

Site: Joma
```{r jomamacroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = joma_adult_df, distributionFamily = negbin(3))

```

## Now fit using poisson

Site: AUCA
```{r macroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = auca_adult_df, distributionFamily = poisson())

```

Site: PRCR
```{r prcrmacroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = prcr_adult_df, distributionFamily = poisson())

```

Site: Rist
```{r ristmacroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = rist_adult_df, distributionFamily = poisson())

```

Site: Demi
```{r demimacroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = demi_adult_df, distributionFamily = poisson())

```

Site: Bowa
```{r bowamacroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = bowa_adult_df, distributionFamily = poisson())

```

Site: Biva
```{r bivamacroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = biva_adult_df, distributionFamily = poisson())

```

Site: Baca
```{r bacamacroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = baca_adult_df, distributionFamily = poisson())

```

Site: Cofr
```{r cofrmacroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = cofr_adult_df, distributionFamily = poisson())

```

Site: Joma
```{r jomamacroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMa(site_df = joma_adult_df, distributionFamily = poisson())

```

## Pooled micro-moth checks

### First negative binomial distributions

Site: AUCA
```{r microNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = auca_adult_df, distributionFamily = negbin(3))

```

Site: PRCR
```{r prcrmicroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = prcr_adult_df, distributionFamily = negbin(3))

```

Site: Rist
```{r ristmicroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = rist_adult_df, distributionFamily = negbin(3))

```

Site: Demi
```{r demimicroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = demi_adult_df, distributionFamily = negbin(3))

```

Site: Bowa
```{r bowamicroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = bowa_adult_df, distributionFamily = negbin(3))

```

Site: Biva
```{r bivamicroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = biva_adult_df, distributionFamily = negbin(3))

```

Site: Baca
```{r bacamicroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = baca_adult_df, distributionFamily = negbin(3))

```

Site: Cofr
```{r cofrmicroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = cofr_adult_df, distributionFamily = negbin(3))

```

Site: Joma
```{r jomamicroNegBin, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = joma_adult_df, distributionFamily = negbin(3))

```

## Now fit using poisson

Site: AUCA
```{r microPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = auca_adult_df, distributionFamily = poisson())

```

Site: PRCR
```{r prcrmicroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = prcr_adult_df, distributionFamily = poisson())

```

Site: Rist
```{r ristmicroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = rist_adult_df, distributionFamily = poisson())

```

Site: Demi
```{r demimicroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = demi_adult_df, distributionFamily = poisson())

```

Site: Bowa
```{r bowamicroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = bowa_adult_df, distributionFamily = poisson())

```

Site: Biva
```{r bivamicroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = biva_adult_df, distributionFamily = poisson())

```

Site: Baca
```{r bacamicroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = baca_adult_df, distributionFamily = poisson())

```

Site: Cofr
```{r cofrmicroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = cofr_adult_df, distributionFamily = poisson())

```

Site: Joma
```{r jomamicroPoisson, echo=FALSE, message=FALSE, warning=FALSE}

generateDHARMaMicro(site_df = joma_adult_df, distributionFamily = poisson())

```